---
title: "Final Project Basic Outline"
author: "Daniel Gao"
date: June 26, 2023
output: html_notebook
---

**Research Question:** 
How do video games genres differ in Esports based on earnings, players?

## Introduction 

The video game industry is growing and one industry that has grown alongside is electronic sports. Electronic sports, otherwise known as Esports, is the competitive aspect of video games where players compete professionally. Either in a team or solo, these players compete against each other in tournaments.

## Data Processing 

#### Data Intake 

This project contains two datasets. Both datasets GeneralEsportData and HistoricalEsportData were taken from [Kaggle.com](https://www.kaggle.com/datasets/rankirsh/esports-earnings). Since both datasets were CSV files, a read.csv() function was used to read the two files. 
```{r}
library(tidyverse)
library(knitr)
library(kableExtra)
library(lubridate)
General <- read.csv("GeneralEsportData.csv") #relative filepath
Historical <- read.csv("HistoricalEsportData.csv") #relative filepath
```
The results of each read function were then assigned to two data frames "General" and "Historical" respectively.

#### Data Wrangling 
 
**General Dataset:**
The General dataset consists of generic data on each game. The dataset contains 5 variables: Game, ReleaseDate, Genre, TotalEarnings, OnlineEarnings, TotalPlayers, and TotalTournaments. There are a total of 631 cases. Each case represents the total earnings of a game since its release date.
```{r}
str(General)
```
**Historical dataset:**
The Historical dataset consists of the earnings per month for every game. The dataset contains 5 variables: Date, Game, Earnings, Players, and Tournaments. There are a total of 8,751 cases. Each case represents the earnings of a game for a month since Jan 1st, 1998 to April 1, 2023. 
```{r}
str(Historical)
```

#### Research Questions
To answer the research question on how video game genres differ in Esports based on earnings, and players, an exploration of the two datasets is needed. 

**What are the genres in Esports? What are each genre's total earnings and total player count?**
```{r}
#for descending, i dont know how to do it in one function
general_desc <- function(results = NULL) { #general_function for multi-use, "results" so I can alter the results table
  General %>%
    select(Genre, TotalEarnings, TotalPlayers) %>%    #dont know if this is needed or not cause of dplyr::select
    group_by(Genre) %>% 
    summarise(TotalEarnings = sum(TotalEarnings), #sum of total earnings per genre
              TotalPlayers = sum(TotalPlayers)) %>% #sum of total players per genre
    arrange(desc(TotalEarnings)) %>%
    mutate(TotalEarnings = paste("$", format(TotalEarnings, big.mark = ","))) %>% #mutate to add $ and comma for money
    mutate(TotalEarnings = gsub("\\s+", "", TotalEarnings)) %>% #there was white space, this mutate deletes them
    mutate(TotalPlayers = paste(format(TotalPlayers, big.mark = ","))) %>% #comma for players
    dplyr::select(Genre, TotalEarnings, TotalPlayers) %>% #makes the result table better for view ability
    rename("Total Earning" = TotalEarnings) %>% #rename
    rename("Total Players" = TotalPlayers) %>% #rename 
    head(results) %>% #where results sits so I can alter the results table
    knitr::kable(caption = "Genres in Esports (Based on Earnings)") %>% 
    kableExtra::kable_styling()
}
#for ascending, i don't know how to do it in one function
general_asc <- function(results = NULL) { #rename
  General %>%
    select(Genre, TotalEarnings, TotalPlayers) %>% #dont know if this is needed or not cause of dplyr::select
    group_by(Genre) %>% 
    summarise(TotalEarnings = sum(TotalEarnings), 
              TotalPlayers = sum(TotalPlayers)) %>% 
    arrange(TotalEarnings) %>%                                                         ##only line this was changed
    mutate(TotalEarnings = paste("$", format(TotalEarnings, big.mark = ","))) %>% 
    mutate(TotalEarnings = gsub("\\s+", "", TotalEarnings)) %>% 
    mutate(TotalPlayers = paste(format(TotalPlayers, big.mark = ","))) %>% 
    dplyr::select(Genre, TotalEarnings, TotalPlayers) %>% 
    rename("Total Earning" = TotalEarnings) %>%
    rename("Total Players" = TotalPlayers) %>% 
    head(results) %>% 
    knitr::kable(caption = "Genres in Esports (Based on Earnings)") %>% 
    kableExtra::kable_styling()
}

general_desc(results = 12)

```
**Result**: In the Genres in Esports table, there consist of a total of 12 total genres. Each genre varies in total earning and popularity based on the total player count for all of each genres tournaments, The genre with the highest total earning is Multiplayer Online Battle Arena with a total earning of $540,062,581.76 and a total player count of 22,955 players.
```{r}
general_desc(results = 1) #highest
general_asc(results = 1) #lowest
```

**Result**: The genre with the lowest total earning is Puzzle Game with a total earning of $74,576.89 with a total player count of 225 players. The differences between the top and bottom genres vary greatly. The difference between total earnings and total players for Multiplayer Online Battle Arena and Puzzle Games is $539,988,004.87 and 22,955, respectively. 

**What were the top 10 games with the highest single month earning? What were the earning, genres, and dates?**

```{r}
HistoricalGeneral <- #joined the Historical and General table cause ymd was needed
  inner_join(Historical, General, by = "Game")

#new dataframe so i can reference old one in case of trouble
hg <- subset(HistoricalGeneral, select = -c(ReleaseDate, TotalEarnings, OnlineEarnings, TotalPlayers, TotalTournaments)) #delete the columns not needed 
hg %>%
  select(Game, Earnings, Genre, Date) %>%
  arrange(desc(Earnings)) %>%
  mutate(Earnings = paste("$", format(Earnings, big.mark = ","))) %>% 
  mutate(Earnings = gsub("\\s+", "", Earnings)) %>% 
  dplyr::select(Game, Earnings, Genre, Date) %>% #makes the tables into nicer tables
  head(10) %>% 
  knitr::kable(caption = "Top 10") %>%
  kableExtra::kable_styling()

```
**Result**: The results of this table show how a single game can dominate in the most earnings within a single month. Dota 2, a game in the Multiplayer Online Battle Arean, appears 7 times in the top 10 games with the highest single-month earnings. Fortnite, a game in the battle royale, appears 2 times out of the top 10. Arena of Valor, a game in the Multiplayer Online Battle Arena genre, appears once. Looking a little deeper into the Multiplayer Online Battle Arena, it appears that Dota 2 represents more than half of the earnings in that genre. 

```{r}
hg %>%
  select(Game, Earnings, Genre) %>%
  group_by(Game, Genre) %>% #group by game and genre is needed 
  summarise(Earnings = sum(Earnings)) %>%
  filter(Genre == "Multiplayer Online Battle Arena") %>%
  arrange(desc(Earnings)) %>%
  mutate(Earnings = paste("$", format(Earnings, big.mark = ","))) %>% 
  mutate(Earnings = gsub("\\s+", "", Earnings)) %>% 
  dplyr::select(Game, Earnings, Genre) %>% 
  head(10) %>% 
  knitr::kable(caption = "Top 10") %>% 
  kableExtra::kable_styling()
```

## Data Visualization
```{r}
testdate <- hg
testdate$Date <- ymd(testdate$Date)
testdate %>% 
  mutate(Year = format(Date, "%Y")) %>%
  group_by(Genre, Year) %>%
  summarise(Earnings = sum(Earnings)/1000000) %>%
  filter(Year %in% c(2010:2023)) %>%
  ggplot(aes(x = Year, y = Earnings, color = Genre, group = Genre)) +
  geom_point() +
  geom_smooth() +
  ggtitle("Genre Earning Per Year (Last 10 Years)") +
  ylab("Earnings (Millions)")
```
This graph depicts the earnings of genres for the last ten years. Throughout the years, there was a rise in earnings from all genres, but genres such as Battle Royale, Multiplayer Online Battle Arena, and First-Person Shooter had an exponential increase in earnings. Compared to the other genres, these three genres make up a notable difference in the earnings of all of the genres.

```{r}
testdate %>% 
  mutate(Year = format(Date, "%Y")) %>%
  group_by(Genre, Year) %>%
  summarise(Players = sum(Players)) %>%
  filter(Year %in% c(2010:2023)) %>%
  ggplot(aes(x = Year, y = Players, color = Genre, fill = Genre)) +
  geom_col() +
  ggtitle("Genre Player Total By Years") +
  ylab("Number of Players")

```
This graph displays the total player of each genre based on years. Although Multiplayer Online Battle Arena has more total earnings, the genre with the consistently highest total players throughout the years is First-Person Shooters. In 2019 and 2023, there can be seen a rise in player count in the Battle Royale genre.

## Conclusion
As the Esports industry grows, so will each of the genres in Esports. But there are differences amongst the genres based on earnings and players. Throughout the years, there can be seen a rise in earnings and popularity in all genres. But one genre stands out above all the other genres in terms of earning. That genre is Multiplayer Online Battle Arena. Within this genre, Dota 2 makes up more than half of the earnings. But alternatively, the genre with the biggest total players is First-Person Shooters. 