---
title: "Initial EDA Basic Outline"
author: "Daniel Gao"
output: html_notebook
---


## Research Question: 
How do video games differ in Esports based on earnings, players, and genre? 

**Introduction**

The video game industry is growing and one industry that has grown alongside is electronic sports. Electronic sports, otherwise known as Esports, is the competitive aspect of video games where players compete professorially. Either in a team, or solo, these players compete against each other in tournaments.  

## Data Processing 

#### Data Intake 

**Importing The Data**

Both datasets "GenralEsportData" and "HistoricalEsportData" were taken from [Kaggle.Com](https://www.kaggle.com/datasets/rankirsh/esports-earnings). Since both datasets were csv files, a read.csv() function were used to read the two files. The result of each read function were then assigned to two dataframes "General" and "Historical" respectively.
```{r}
General <- read.csv("GeneralEsportData.csv")
Historical <- read.csv("HistoricalEsportData.csv")
```

#### Data Wrangling  

**General Dataset**:

The General dataset consist of generic data on each game. The dataset contains 5 variables: Game, ReleaseDate, Genre,
TotalEarnings, OnlineEarnings, TotalPlayers, and TotalTournaments. There are a total of 631 cases. These cases represent a single distinct game. 
```{r}
str(General)
summary(General)
```

**A little more about the General dataset**:

There are no null values in this data.
```{r}
colSums(is.na(General))
```
According to the table and graph, the genre with the highest earning is Multiplayer Online Battle Arena with a total earning of $540,062,582. The genre with the second high earning is First-Pesron Shooter with a total earning of $380,906,784. 
```{r}
General %>% 
  group_by(Genre) %>%
  summarise(TotalEarnings = sum(TotalEarnings)) %>%
  arrange(desc(TotalEarnings)) %>%
  head()

General %>% 
  ggplot(aes(x = Genre, y = TotalEarnings/1000000)) +
  geom_boxplot() +
  ylab("Total Earnings (Millions)")
```
**Historical dataset**
The Historical dataset consist of the earnings per month for every game. The dataset contains 5 variables: Date, Game, Earnings, Players, and Tournaments. There are a total of 8751 cases with each case representing the earning of a game for that month. 
```{r}
str(Historical)
summary(Historical)
```
***A little more about the Historical dataset***
There are no null values in this dataset.
```{r}
colSums(is.na(Historical))
```
According to the table and graph, the game with the highest earnings for all the tournaments was Dota 2 with a total earning of $315,065,846. The game with second highest earning is Fortnite $155,314,332. 
```{r}
Historical %>%
  group_by(Game) %>%
  summarise(Earnings = sum(Earnings)) %>%
  arrange(desc(Earnings)) %>%
  head()

Historical %>% 
  ggplot(aes(x = Game, y = Earnings)) +
  geom_boxplot()
```
**How was the data colletect**: The creator, Ran.Kirsh, obtain this data from [Esports Earnings](https://www.esportsearnings.com/tournaments). Esports Earning stated, "Sources include news articles, forum posts, live report threads, interviews, official statements, reliable databases, VODs and other publicly-accessable sources that preserve "historical" information.


## Data Visualization

#### Plot 1

Include R code to make one data visualization (I would encourage you to use 3 or more variables). Include a brief written description of what your graph shows. Is this graph helpful in answering your research question? If it is helpful, will you need to add/modify anything in this graph in the final report? 

#### Plot 2

Include R code to make one data visualization (I would encourage you to use 3 or more variables). Include a brief written description of what your graph shows. Is this graph helpful in answering your research question? If it is helpful, will you need to add/modify anything in this graph in the final report? 


## Conclusion

Include your initial intuition as to what the answer to your research question is using the data sets and visualizations you created. 