---
title: "Initial EDA Basic Outline"
author: "Daniel Gao"
output: html_notebook
---


## Research Question: 
How do video games differ in Esports based on earnings, players, and genre? 

**Introduction**

The video game industry is growing and one industry that has grown alongside is electronic sports. Electronic sports, otherwise known as Esports, is the competitive aspect of video games where players compete professorially. Either in a team or solo, these players compete against each other in tournaments.  

## Data Processing 

#### Data Intake 

**Importing The Data**

Both datasets "GenralEsportData" and "HistoricalEsportData" were taken from [Kaggle.com](https://www.kaggle.com/datasets/rankirsh/esports-earnings). Since both datasets were CSV files, a read.csv() function was used to read the two files. The results of each read function were then assigned to two data frames "General" and "Historical" respectively.
```{r}
General <- read.csv("GeneralEsportData.csv")
Historical <- read.csv("HistoricalEsportData.csv")
library(tidyverse)
```

#### Data Wrangling  

**General Dataset**:

The General dataset consists of generic data on each game. The dataset contains 5 variables: Game, ReleaseDate, Genre, TotalEarnings, OnlineEarnings, TotalPlayers, and TotalTournaments. There are a total of 631 cases. Each case represents a single distinct game. 
```{r}
str(General)
summary(General)
```

**A little more about the General dataset**:

There are no null values in this data.
```{r}
colSums(is.na(General))
```
According to the table and graph, the genre with the highest earning is Multiplayer Online Battle Arena with a total earning of $540,062,582. The genre with the second high earning is First-Pesron Shooter with a total earning of $380,906,784. 
```{r}
General %>% 
  group_by(Genre) %>%
  summarise(TotalEarnings = sum(TotalEarnings)) %>%
  arrange(desc(TotalEarnings)) %>%
  head()

General %>% 
  group_by(Genre) %>%
  summarise(TotalEarnings = sum(TotalEarnings)) %>%
  arrange(desc(TotalEarnings)) %>%
  ggplot(aes(x = Genre, y = TotalEarnings/1000000)) +
  geom_col(fill = "blue") +  
  coord_flip() +
  ylab("Total Earnings (Millions)") +
  ggtitle("Tournament Earnings For Each Genre")
```
**Historical dataset**
The Historical dataset consists of the earnings per month for every game. The dataset contains 5 variables: Date, Game, Earnings, Players, and Tournaments. There are a total of 8751 cases. Each case represents the earning of a game for that month. 
```{r}
str(Historical)
```
***A little more about the Historical dataset***
There are no null values in this dataset.
```{r}
colSums(is.na(Historical))
```
According to the table and graph, the game with the highest earnings for all the tournaments was Dota 2 with a total earning of $315,065,846. The game with the second highest earning is Fortnite $155,314,332. 
```{r}
Historical %>%
  group_by(Game) %>%
  summarise(Earnings = sum(Earnings)) %>%
  arrange(desc(Earnings)) %>%
  head()

Historical %>%
  group_by(Game) %>%
  summarise(Earnings = sum(Earnings)) %>%
  arrange(desc(Earnings)) %>%
  head(10) %>%
  ggplot(aes(x = Game, y = Earnings / 1000000)) +
  geom_col(fill = "blue") +
  coord_flip() +
  ggtitle("Top 10 Tournament Earnings ") +
  ylab("Earnings (Millions)")
```
**How the data was colletecd**: The creator, Ran.Kirsh, obtain this data from [Esports Earnings](https://www.esportsearnings.com/tournaments). Esports Earning stated, "Sources include news articles, forum posts, live report threads, interviews, official statements, reliable databases, VODs, and other publicly-accessible sources that preserve "historical" information.

## Data Visualization

#### Plot 1
This plot displays the top 10 games with the highest total players in all of their tournaments combined. The game with the highest total of players is Counter Strike Global Offensive. The game with the second-highest total player count is League of Legends. This graph does help in answering my research question, but I will need to modify this graph. The x-axis and y-axis do not are not refined.  
```{r}
General %>% 
  arrange(desc(TotalPlayers)) %>%
  head(10) %>%
  ggplot(aes(x = Game, y = TotalPlayers, color = Genre)) +
  geom_point(shape = "circle", size = 1) +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Top 10 Games (Total Tournament Play Count)")
```
#### Plot 2
This graph displays the top 10 games with the highest single tournament earning. Dota 2 held 5 tournaments with the highest earnings and Fortnite held 1 tournament with the highest earnings. This graph does help in answering my research question, but there are several things I would want to change from this graph. The first would be the y scale as the scale shows scientific notation.
```{r}
Historical %>%
  arrange(desc(Earnings)) %>%
  head(10) %>%
  ggplot(aes(x = Game, y = Earnings / 1000000, color = Date)) +
  geom_point(shape = "circle", size = 1) +
  ylab("Earnings (Millions)") +
  ggtitle("Top 10 Tournaments Earnings (Game) ")
```

## Conclusion

The answer to my research question is a comparison between earnings, player, and genre. I have been diving into the differences between games in the Esports scene based on earnings, players, and genre. But I have not been able to fully utilize the two datasets in terms of joining the two datasets. I would like the create graph that displays the join also. The graphs are another objective I would like to work on. There some graphs that are not fully refined.
